### 2022-04-24 paper/Chapters/Background.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

On the corporate side of computing, the solution for portability has been the thin client.
Rather than integrating the power of a desktop to a portable machine, a thin client will connect to a powerful server and provide a thin interface to the user.
This allows the user to utilize the power of the host machine from a cheap computer that usually doesn't have enough power to be used as a standalone computer.
Because of it's low power, the thin client cannot be used as a standalone computer, but it's low cost and portability make it a great solution for a corporate user.
While these clients work well for simple interfacing tasks such as data entry or system management, they are not designed to be used for complex tasks that require quick response times or graphical fidelity.

>>>>

On the corporate side of computing, the solution for portability has been the thin client.
Rather than integrating the power of a desktop to a portable machine, a thin client will connect to a powerful server and provide a thin interface to the user.
This allows the user to utilize the power of the host machine from a cheap computer that usually doesn't have enough power to be used as a standalone computer.
Because of this, the thin client cannot be used as a standalone computer, but it's low cost and portability make it a great solution for a corporate user.
While these clients work well for simple interfacing tasks such as data entry or system management, they are not designed to be used for complex tasks that require quick response times or graphical fidelity.

### 2022-04-24 paper/Chapters/Conclusion.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

This thesis has proved that it is possible to produce a hardware and software solution that is capable of leveraging the power of a desktop computer remotely without sacrificing the user's experience.
It improves upon existing solutions by continuing to offer a performant and high-quality stream of the host's screen and audio to the user even as the host is running highly demanding applications.
Each of the questions posed in Section \ref{sec:ResearchQuestions} can be answered individually.

@@

\noindent
This thesis project is able to run on embedded Single Board Computers and hardware available for purchase today.
While this project built a custom PCB for the Raspberry Pi Compute Module 4 (Discussed in section \ref{sec:ChoosingParts}), the software could also run on consumer-grade Raspberry Pi models as well.

>>>>

This thesis has proved that it is possible to produce a hardware and software solution that is capable of leveraging the power of a desktop computer remotely without sacrificing the user's experience.
It improves upon existing solutions by continuing to offer a performant and high-quality stream of the host's screen to the user even as the host is running highly demanding applications.
Each of the questions posed in Section \ref{sec:ResearchQuestions} can be answered individually.

@@

\noindent
This thesis project is able to run on embedded Single Board Computers and hardware available for purchase today.
While this project built a custom PCB for the Raspberry Pi Compute Module 4 (Discussed in section \ref{sec:ChoosingParts}), the software could also run on consumer-grade main-line Raspberry Pi models as well.

### 2022-04-24 paper/Chapters/Evaluation.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

In order to evaluate the performance of the forked version of Moonlight developed for this paper, it must be tested against the other existing solutions.
The testing methodology employed to ensure the accuracy of the testing is described in Section \ref{sec:TestingMethodology}.
Then, the different results are evaluated and analyzed in Sections \ref{sec:ResponsivityAndLatency} and \ref{sec:Quality}.
Finally, the results are summarized and verdicts on each of the solutions' applicability to the research questions will be presented in Section \ref{sec:EvaluationSummary}.

@@

\begin{enumerate}
  \item Both the host and client computers are fully rebooted to ensure no other application is running.
  \item The host computer starts up a single browser window with the web page described in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
  \item If the test calls for the CPU or the GPU to be stressed, the application \enquote{Blender} is started on the host machine, and a benchmark is performed for the CPU or the GPU.
        This benchmark will run for longer than the duration of the test, and will keep the CPU or GPU at 100\% load for the entire duration of the benchmark.
  \item Both devices start up their respective data collection tool that will record key-strokes and changes in the screen's color.
  \item The client machine connects to the host computer using the application being tested.
  \item The tester presses the \enquote{=} to begin the automated testing process.
        Due to the wide variance in what the user can do using the software, it is nearly impossible to test real world scenarios.
        Instead, the automated test simulates an active working environment by sending 400 keystrokes with a 500ms delay between inputs coupled with updating the entire screen for each keystroke to mimic high-intensity applications.
  \item Once the trial is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
\end{enumerate}

@@

\noindent
Once all the data is collected, the data from the host and client machines for a single trial are organized into a single CSV file for processing.
In order to ensure that the data is comparable, all extraneous factors must be removed from the data.
Namely, the network latency and difference in computer clocks must be removed.
By keeping the host and client computers attached to the same network through ethernet, the network latency is minimized.
But the task of keeping two computer's internal clocks synchronized is not trivial.
Even after instructing the operating system to synchronize its clock with an internet time server, the two computers may still report different times.
In some cases, the difference between the host and client computer's system time was noticed to be upwards of a full second after resynchronizing both clocks \cite{time.is}.
This is partially accounted for in step \emph{6} of the above process, and the remaining inaccuracy can now be removed by subtracting the time difference between the when the client presses the button and the time the host executes that command.
This ensures that any inaccuracies in each computer's internal clock will not impact the data.

After the data is cleaned up, the data can be processed by matching each key-stroke and change in the screen's color detected by the client and host machines and calculating how much time elapsed between the events.
This reveals the amount of delay the user experiences between when they press a key and when the host executes the command, and between when they press a key and can visually see the response.

@@

The first factor in determining the performance of the application is to take a look at how responsive it is to use.
Responsivity is defined as how quick the application responds to inputs and how quickly the user can see the results of their inputs.
This is important for general user experience and applications where continuous input is required.
This is also what is generally referred to when determining whether an application has a positive user experience, since a program that can take seconds to respond to the user begets lower productivity and a more frustrating experience.
To analyze this, two points of data are recorded: First, the amount of time between when the client records a keystroke and when it is actually inputted on the host machine, and secondly, the amount of time between the host machine updating their screen and when the client actually displays that update.
The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different programs may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
Each of the following figures shows data over one trial for each category.
Each trial consists of the client machine programmatically sending 400 keystrokes with 500ms between each input.
Limitations in this data is discussed in Section \ref{sec:ConclusionLimitations}.
Figure \ref{fig:InputDelay} shows the amount of time between the client and the host registering the keystroke under the different test conditions.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        boxplot/draw direction=y,
        ylabel={Time (ms)},
        scale only axis,
        height=8cm,
        width=\textwidth,
        ymin=0,ymax=200,
        cycle list={{green!70!black},{red},{blue},{black}},
        legend style={
            legend pos=outer north east,
            font=\small
          },
        legend cell align=left,
        legend image code/.code={%
            \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
          },
        boxplot={
            % Three boxplots for each column
            draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*fpumod(\plotnumofactualtype,4)},
            % Each plot takes up a 15% of the column
            box extend=0.15,
          },
        % 1 unit in x controls the width:
        x=2cm,
        % ... and it means that we should describe intervals:
        xtick={0,1,2,...,10},
        x tick label as interval,
        xticklabels={%
            {Forked Moonlight},%
            {CRD},%
            {RDP},%
            {VNC},%
          },
        x tick label style={
            text width=2.5cm,
            align=center
          },
      ]

@@

      \legend{Ideal,CPU stressed,GPU stressed, Both stressed}
    \end{axis}
  \end{tikzpicture}
  \caption[Input Delay Data]{Time for host to press key from client. n=400.}
  \label{fig:InputDelay}
\end{figure}

Each application does a good job sending data from the client to the host in a timely manner, with the majority of keystrokes being realized within a tenth of a second of being pressed on the client machine.
The forked Moonlight implementation keeps a latency under 20ms regardless of the load of the host machine, with more than 75\% of all keystrokes being realized within 10ms.
CRD proves to be a CPU limited application, with latency while idle or under GPU load also staying below 20ms, but climbing up above 50ms when under CPU load.
This is important to keep in mind when thinking about what sort of situations the host machine will be in while being accessed.
If the user is going to be rendering video remotely, a GPU intensive process, the latency of the stream wont be impacted, but if the user is going to be running a CPU intensive process, the latency can be expected to be negatively impacted.
RDP has a similar results, with extremely low latency while idle or under GPU load, and upwards of 40ms while under CPU load.
VNC has the highest variance, with no discernable affinity to any kind of load.
This lines up with it's intended use described in Section \ref{subsec:VirtualNetworkComputing}, being the simplest and most straightforward method of remote desktop control without much focus on performance or security.
That's not so say VNC is unusable or even difficult to use; latency of 140ms is still usable in all but the most precise situations.
Though this is only one metric to consider, the other is the amount of time it take for the client to display the results of their input.
This is recorded as the amount of time it takes between the host machine changing the color of it's screen and the client rendering the change.
As noted before, the color of the host's screen changes with each input sent by the client, which occurs every 500ms for 400 keystrokes.
Figure \ref{fig:ColorDelay} shows the amount of time for each application under each condition.

@@

      \legend{Ideal,CPU stressed,GPU stressed,Both stressed}
    \end{axis}
  \end{tikzpicture}
  \caption[Color Delay Data]{Time for client to render update to the screen. n=400.}
  \label{fig:ColorDelay}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        ylabel={Color changes not rendered (\%)},
        scale only axis,
        height=8cm,
        width=\textwidth,
        enlarge x limits=0.25,
        ybar=2*\pgflinewidth,
        bar width=14pt,
        ymin=0,ymax=100,
        cycle list={{green!70!black},{red},{blue},{black}},
        legend style={
            legend pos=outer north east,
            font=\small
          },
        legend cell align=left,
        legend image code/.code={%
            \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
          },
        % 1 unit in x controls the width:
        x=2cm,
        % ... and it means that we should describe intervals:
        xtick={0,1,2,...,10},
        % x tick label as interval,
        xticklabels={%
            {Forked Moonlight},%
            {CRD},%
            {RDP},%
            {VNC},%
          },
        x tick label style={
            text width=2.5cm,
            align=center
          },
      ]

@@

Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
Each application will be assessed individually, and then compared in Section \ref{sec:EvaluationSummary}
When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
Once the streaming session concludes, the statistics are collated into human-readable numbers and presented in the console.
What follows is an example output from one of the test streaming sessions:

\begin{lstlisting}[style=plaintext,title=Statistics generated after a streaming session]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 62.18 FPS
Decoding frame rate: 62.18 FPS
Rendering frame rate: 62.18 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.27 ms
Average rendering time (including monitor V-sync latency): 0.06 ms

Total dropped frames: 0
\end{lstlisting}

@@

@@

@@

@@

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}{1\textwidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/realworld/kicad}
  \end{subfigure}
  \begin{subfigure}{1\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{Figures/realworld/kicadstats}
  \end{subfigure}
  \caption[Streaming Electronics Design software]{Working on the PCB for this project using KiCad.}
  \label{fig:RealWorldKicad}
\end{figure}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming KiCad (Figure \ref{fig:RealWorldKicad})]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 61.23 FPS
Decoding frame rate: 61.23 FPS
Rendering frame rate: 61.23 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.24 ms
Average rendering time (including monitor V-sync latency): 0.06 ms

Total dropped frames: 0
\end{lstlisting}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while training an AI (Figure \ref{fig:RealWorldAI})]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 62.05 FPS
Decoding frame rate: 62.05 FPS
Rendering frame rate: 62.05 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.26 ms
Average rendering time (including monitor V-sync latency): 0.05 ms

Total dropped frames: 0
\end{lstlisting}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming Blender (Figure \ref{fig:RealWorldBlender})]
  Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 57.30 FPS
Decoding frame rate: 57.30 FPS
Rendering frame rate: 57.30 FPS
Frames dropped by your network connection: 0.06%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.70 ms
Average rendering time (including monitor V-sync latency): 0.05 ms

Total dropped frames: 16
\end{lstlisting}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming the development of a video game level (Figure \ref{fig:RealWorldGameDev})]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 61.93 FPS
Decoding frame rate: 61.93 FPS
Rendering frame rate: 61.93 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.27 ms
Average rendering time (including monitor V-sync latency): 0.06 ms

Total dropped frames: 0
\end{lstlisting}

>>>>

In order to evaluate the performance of the forked version of Moonlight developed for this paper, it must be tested against the other existing solutions.
The testing methodology employed to ensure the accuracy of the testing is described in Section \ref{sec:TestingMethodology}.
Then, the responsivity and latency of each solution is evaluated in Section \ref{sec:ResponsivityLatency}, and the quality of each solution's stream is evaluated in Section \ref{sec:Quality}.
Then, the forked version of Moonlight is evaluated in real world scenarios in Section \ref{sec:RealWorldTesting}.
Finally, the results are summarized and verdicts on each of the solutions' applicability to the research questions will be presented in Section \ref{sec:EvaluationSummary}.

@@

\begin{enumerate}
  \item Both the host and client computers are fully rebooted to ensure no other application is running.
  \item The host computer starts up a single browser window with the web page described in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
  \item If the test calls for the CPU or the GPU to be stressed, the application \enquote{Blender} is started on the host machine, and a benchmark is performed for the CPU or the GPU.
        This benchmark will run for longer than the duration of the test, and will keep the CPU and/or GPU at 100\% load for the entire duration of the benchmark.
  \item Both devices start up their respective data collection tool that will record keystrokes and changes in the screen's color.
  \item The client machine connects to the host computer using the application being tested.
  \item The tester presses the \enquote{=} key on the client machine to begin the automated testing process.
        Due to the wide variance in what the user can do using the software, it is nearly impossible to test every possible real world scenario.
        Instead, the automated test simulates an active working environment by sending a repeatable series of keystrokes that cause the entire screen to change colors to mimic high-intensity applications.
        For consistent testing, 400 keystrokes are automatically sent with a 500ms delay between inputs.
  \item Once the trial is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
\end{enumerate}

@@

\noindent
Once all the data is collected, the data from the host and client machines for a single trial are organized into a single CSV file for processing.
In order to ensure that the data is comparable, all extraneous factors must be removed from the data.
Namely, the network latency and difference in computer clocks must be removed.
By keeping the host and client computers attached to the same network through ethernet, the network latency is minimized.
But the task of keeping two computer's internal clocks synchronized is not trivial.
Even after instructing the operating system to synchronize its clock with an internet time server, the two computers may still report different times.
In some cases, the difference between the host and client computer's system time was noticed to be upwards of a full second after resynchronizing both clocks \cite{time.is}.
This is partially accounted for in step \emph{6} of the above process, and the remaining inaccuracy can now be removed by subtracting the time difference between the when the client presses the button and the time the host executes that command.
It is important to not subtract the entire round trip time of the input, as that is what is being testing in the following section, but rather just the time between when the client sends the input and when the host receives it.
This ensures that any inaccuracies in each computer's internal clock will not impact the data.

After the data is cleaned up, the data can be processed by matching each keystroke and change in the screen's color detected by the client and host machines and calculating how much time elapsed between the events.
This reveals the amount of delay the user experiences between when they press a key and when the host executes the command, and between when they press a key and can visually see the response.

@@

The first factor in determining the performance of the application is to take a look at how responsive it is to use.
Responsivity is defined as how quick the application responds to inputs and how quickly the user can see the results of their inputs.
This is important for general user experience and applications where continuous input is required.
This is also what is generally referred to when determining whether an application has a positive user experience, since a program that can take seconds to respond to the user begets lower productivity and a more frustrating experience.
To analyze this, two points of data are recorded: First, the amount of time between when the client records a keystroke and when it is actually inputted on the host machine, and secondly, the amount of time between the host machine updating their screen and when the client actually displays that update.
The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different solutions may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
This means that as the network delay is removed from the data, all that is left is the time that it takes the software on the host's machine to actually execute the input.
Each of the following figures shows data over one trial for each category.
Each trial consists of the client machine programmatically sending 400 keystrokes with 500ms between each input.
Limitations in this data is discussed in Section \ref{sec:ConclusionLimitations}.
\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        boxplot/draw direction=y,
        ylabel={Time (ms)},
        scale only axis,
        height=8cm,
        width=\textwidth,
        ymin=0,ymax=200,
        cycle list={{green!70!black},{red},{blue},{black}},
        legend style={
            legend pos=outer north east,
            font=\small
          },
        legend cell align=left,
        legend image code/.code={%
            \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
          },
        boxplot={
            % Three boxplots for each column
            draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*fpumod(\plotnumofactualtype,4)},
            % Each plot takes up a 15% of the column
            box extend=0.15,
          },
        % 1 unit in x controls the width:
        x=2cm,
        % ... and it means that we should describe intervals:
        xtick={0,1,2,...,10},
        x tick label as interval,
        xticklabels={%
            {Forked Moonlight},%
            {CRD},%
            {RDP},%
            {VNC},%
          },
        x tick label style={
            text width=2.5cm,
            align=center
          },
      ]

@@

      \legend{Ideal,CPU stressed,GPU stressed, Both stressed}
    \end{axis}
  \end{tikzpicture}
  \caption[Input Delay Data]{Time for host to press key from client. n=400.}
  \label{fig:InputDelay}
\end{figure}
Figure \ref{fig:InputDelay} shows the amount of time between the client and the host registering the keystroke under the different test conditions.

Each application does a good job sending data from the client to the host in a timely manner, with the majority of keystrokes being realized within a tenth of a second of being pressed on the client machine.
The forked Moonlight implementation keeps a latency under 20ms regardless of the load of the host machine, with more than 75\% of all keystrokes being realized within 10ms.
CRD proves to be a CPU limited application, with latency while idle or under GPU load also staying below 20ms, but climbing up above 50ms when under CPU load.
This is important to keep in mind when thinking about what sort of situations the host machine will be in while being accessed.
If the user is going to be rendering video remotely, a GPU intensive process, the latency of the stream wont be impacted, but if the user is going to be running a CPU intensive process, the latency can be expected to be negatively impacted.
RDP has a similar results, with extremely low latency while idle or under GPU load, and upwards of 40ms while under CPU load.
VNC has the highest variance, with no discernable affinity to any kind of load.
This lines up with it's intended use described in Section \ref{subsec:VirtualNetworkComputing}, being the simplest and most straightforward method of remote desktop control without much focus on performance or security.
That's not so say VNC is unusable or even difficult to use; latency of 140ms is still usable in all but the most precise situations.

Though this is only one metric to consider, the other is the amount of time it take for the client to display the results of their input.
This is recorded as the amount of time it takes between the host machine changing the color of it's screen and the client rendering the change.
As noted before, the color of the host's screen changes with each input sent by the client, which occurs every 500ms for 400 keystrokes.
Figure \ref{fig:ColorDelay} shows the amount of time for each application under each condition.

@@

      \legend{Ideal,CPU stressed,GPU stressed,Both stressed}
    \end{axis}
  \end{tikzpicture}
  \caption[Color Delay Data]{Time for client to render update to the screen. Values over 250ms push the usability of the system. n=400.}
  \label{fig:ColorDelay}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        ylabel={Color changes not rendered (\%)},
        scale only axis,
        height=8cm,
        width=\textwidth,
        enlarge x limits=0.25,
        ybar=2*\pgflinewidth,
        bar width=14pt,
        ymin=0,ymax=100,
        cycle list={{green!70!black},{red},{blue},{black}},
        legend style={
            legend pos=outer north east,
            font=\small
          },
        legend cell align=left,
        legend image code/.code={%
            \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
          },
        % 1 unit in x controls the width:
        x=2cm,
        % ... and it means that we should describe intervals:
        xtick={0,1,2,...,10},
        % x tick label as interval,
        xticklabels={%
            {Forked Moonlight},%
            {CRD},%
            {RDP},%
            {VNC},%
          },
        x tick label style={
            text width=2.5cm,
            align=center
          },
      ]

@@

Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
Each application will be assessed individually, and then compared in Section \ref{sec:EvaluationSummary}.
Due to the closed nature of some of the protocols, only limited amounts of data could be extracted and compared.
This prevents direct comparisons, but can still contribute to the understanding of the overall performance of the solution.

\subsection{Moonlight}\label{subsec:QualityMoonlight}

When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
Once the streaming session concludes, the statistics are collated into human-readable numbers and presented in the console.
What follows is an example output from one of the test streaming sessions:

\begin{lstlisting}[style=plaintext,title=Statistics generated after a streaming session]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 62.18 FPS
Decoding frame rate: 62.18 FPS
Rendering frame rate: 62.18 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.27 ms
Average rendering time (including monitor V-sync latency): 0.06 ms
Total dropped frames: 0
\end{lstlisting}

@@

\subsection{Chrome Remote Desktop}\label{subsec:QualityCRD}

@@

\subsection{Remote Desktop Protocol}\label{subsec:QualityRDP}

@@

\subsection{Virtual Network Computing}\label{subsec:QualityVNC}

@@

\begin{figure}[!b]
  \centering
  \begin{subfigure}{1\textwidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/realworld/kicad}
  \end{subfigure}
  \begin{subfigure}{1\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{Figures/realworld/kicadstats}
  \end{subfigure}
  \caption[Streaming Electronics Design software]{Working on the PCB for this thesis using KiCad.}
  \label{fig:RealWorldKicad}
\end{figure}

@@

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming KiCad (Figure \ref{fig:RealWorldKicad})]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 61.23 FPS
Decoding frame rate: 61.23 FPS
Rendering frame rate: 61.23 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.24 ms
Average rendering time (including monitor V-sync latency): 0.06 ms
Total dropped frames: 0
\end{lstlisting}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while training an AI (Figure \ref{fig:RealWorldAI})]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 62.05 FPS
Decoding frame rate: 62.05 FPS
Rendering frame rate: 62.05 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.26 ms
Average rendering time (including monitor V-sync latency): 0.05 ms
Total dropped frames: 0
\end{lstlisting}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming Blender (Figure \ref{fig:RealWorldBlender})]
  Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 57.30 FPS
Decoding frame rate: 57.30 FPS
Rendering frame rate: 57.30 FPS
Frames dropped by your network connection: 0.06%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.70 ms
Average rendering time (including monitor V-sync latency): 0.05 ms
Total dropped frames: 16
\end{lstlisting}

@@

\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming the development of a video game level (Figure \ref{fig:RealWorldGameDev})]
Global Video Stats
----------------------------------------------------------
Incoming frame rate from network: 61.93 FPS
Decoding frame rate: 61.93 FPS
Rendering frame rate: 61.93 FPS
Frames dropped by your network connection: 0.00%
Average network latency: 1 ms (variance: 0 ms)
Average decoding time: 4.27 ms
Average rendering time (including monitor V-sync latency): 0.06 ms
Total dropped frames: 0
\end{lstlisting}

### 2022-04-24 paper/Chapters/Hardware.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

This section will explain the different constraints taken into account while developing the hardware portion of this project.
First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).

Due to the lack of preliminary studies or research into the topic of target audience for this project, the requirements laid out in this section are based on the author's own requirements and preferences.
The requirements are constructed with the target user being someone who already has a working desktop computer and wishes to leverage it's power remotely without sacrificing the user experience.

@@

When starting to choose parts for the hardware, there are many difficult questions to face.
First and foremost, the hardware had to be built with the time and resources available for this project.
Specifically, building a Single Board Computer from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
Boasting an incredibly small size and streamlined I/O, there only way to communicate with the Compute Module was through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
With no other pin outs for anything such as power, HDMI, or USB, everything had to be built on a custom Printed Circuit Board (PCB).

>>>>

This section will explain the different constraints taken into account while developing the hardware portion of this thesis.
First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).

Since no specific studies regarding possible users of the system were conducted, the requirements laid out in this section are based on the author's own requirements and preferences.
The requirements are constructed with the target user being someone who already has a working desktop computer and wishes to leverage it's power remotely without sacrificing the user experience.

@@

When starting to choose parts for the hardware, there are many difficult questions to face.
First and foremost, the hardware had to be built with the time and resources available for this project.
Specifically, building a Single Board Computer from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
As this perfectly lines up with the requirements for the project, this thesis will be built using the CM4.
However, due to it's small size and deeply embedded nature, the only way to communicate with the Compute Module is through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
With no other pin outs for anything such as power, HDMI, or USB, the rest of the hardware must be built on a custom Printed Circuit Board (PCB) that tailors the CM4 for this thesis.

### 2022-04-24 paper/Chapters/Introduction.tex
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.
Implement Director feedback
Everything except for summaries for each section and a last look-over

\chapter{Introduction} % Main chapter title

\label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

@@

The primary goal of this thesis is to create a portable hardware and software solution that can be used to control a desktop computer remotely while leveraging all of it's power.
There are existing software solutions that give a user remote control of their desktop computer from a different location, but many of them fail to provide a high-quality and performant solution to working with highly demanding programs that put the desktop computer under intense load.
Once attempting to run an intensive application such as rendering video, training Artificial Intelligence, or developing a video game, existing solutions fail to keep up.

>>>>

\chapter{Introduction}

\label{Chapter1}

@@

The primary goal of this thesis is to create a portable hardware and software solution that can be used to control a desktop computer remotely while leveraging all of it's power.
There are existing software solutions that give a user remote control of their desktop computer from a different location, which will be introduced in Chapter \ref{Chapter3}, but such solutions may not provide a performant and high-quality  experience for the user.
These solutions will be tested while running highly demanding applications such as rendering animation, training Artificial Intelligence models, or developing video games to see if they are capable of working in such situations.

### 2022-04-24 paper/Chapters/Software.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

By taking advantage of the Moonlight library briefly described in Section \ref{subsec:NvidiaShieldAndGameStream}, the previously described process is translated  into a usable programming interface.
To accomplish this, a fork of the Moonlight project is created so the GameStream protocol can be leveraged in this thesis.
With the fork, much of the code required to create the software solution is already provided, all that is left is to tweak the code to ease development for the \emph{ARM} architecture and to build a suite of testing tools that hook into Moonlight to extract statistics.

@@

In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
To do this, two sets of data will be collected: First, data is recorded directly from the forked version of moonlight that logs statistics such as the number of frames received, dropped, or rendered, and data that is recorded from a separate application that logs the amount of time it takes for the client machine to receive updates after it sends data to the host machine.
The first set of data is important to track the raw performance of the application, with statistics like Frames Per Second (FPS) and decoding time which are difficult to record otherwise.
These statistics provide insight into how performant the application is from a technical perspective.
The following is a structure that was developed for this thesis to hold the statistics recorded over a session of connecting to a host computer.
Corresponding code was written throughout the fork of the Moonlight application to populate this structure while a streaming session is running.
The data recorded is discussed in Chapter \ref{Chapter6}.

>>>>

By taking advantage of the Moonlight library briefly described in Section \ref{subsec:NvidiaShieldAndGameStream}, the previously described process is translated  into a usable programming interface.
To accomplish this, a fork of the Moonlight project is created so the GameStream protocol can be leveraged in this thesis.
With the fork, much of the code required to create the software solution is already provided, the next step is to build on top of the existing code to ease development for the \emph{ARM} architecture and to write a suite of testing tools that hook into Moonlight to extract statistics.

@@

In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
To do this, two sets of data will be collected: First, data is recorded directly from the forked version of moonlight that logs statistics such as the number of frames received, dropped, or rendered, and data that is recorded from a separate application that logs the amount of time it takes for the client machine to receive updates after it sends data to the host machine.
The first set of data is important to track the raw performance of the application, with statistics like Frames Per Second (FPS) and decoding time which are difficult to record otherwise.
These statistics provide insight into how performant the application is from a technical perspective.
The following is a structure that was developed for this thesis to hold the statistics recorded over a session of connecting to a host computer.
Corresponding code was written throughout the fork of the Moonlight application to populate this structure while a streaming session is running.
The data recorded is evaluated in Chapter \ref{Chapter6}.
